---
title: "ESM 244 Lab 7 - Spatial Data"
author: "Deanna Elliott"
date: "2/17/2022"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(broom)
library(sf)
library(tmap)
library(gstat)
library(stars)
```

# Part 1: Spatial data wrangling, visualization


### A. California county outlines (polygons)

```{r}

ca_counties_sf <- read_sf(here("data", "ca_counties", 
                               "CA_Counties_TIGER2016.shp")) 

ca_subset_sf <- ca_counties_sf %>% 
  janitor::clean_names() %>% 
  select(county_name = name, land_area = aland) #sticky geography - keeps spatial data attached, despite not selecting

ca_counties_df <- ca_counties_sf %>% 
  as.data.frame() %>% 
  select(-geometry)

```

### Examine and set CRS

```{r}

ca_subset_sf %>% st_crs()

ca_subset_sf %>%  raster::crs()
```

```{r}

ggplot(data = ca_subset_sf) +
  geom_sf(aes(fill = land_area), color = 'white', size = 0.1) +
  theme_void() +
  scale_fill_gradientn(colors = c('cyan', 'blue', 'purple')) +
  labs(fill = "Land Area")
```

### B. Invasive red sesbania records (spatial points)

```{r}

sesbania_sf <- read_sf(here("data", "red_sesbania", "ds80.shp")) %>% 
  janitor::clean_names()

# check CRS:
sesbania_sf %>% st_crs()

# if you know EPSG code
sesbania_3857_sf <- st_transform(sesbania_sf, 3857)
# if you don't know EPSG
sesbania_3857_2_sf <- st_transform(sesbania_sf, st_crs(ca_subset_sf))

#check CRS
sesbania_3857_sf %>%  st_crs()
```

### Plot then together!

```{r}

ggplot() +
  geom_sf(data = ca_subset_sf) +
  geom_sf(data = sesbania_3857_sf, size = 1, color = 'red')
```

### Wrangle to fnd sesbania observations per county

```{r}

ca_ses_sf <- ca_subset_sf %>% 
  st_join(sesbania_3857_sf)

ses_counts_sf <- ca_ses_sf %>% 
  group_by(county_name) %>% 
  summarize(n_records = sum(!is.na(id)))
```

```{r}

ggplot(data = ses_counts_sf) +
  geom_sf(aes(fill = n_records), color = 'white', size = 0.1) +
  scale_fill_gradientn(colors = c('lightgrey', 'orange', 'red')) +
  theme_void() +
  labs(fill= 'Number of S. punicea records')
```

### Filter to just occurances in Solano County

```{r}

# subset sesbania based on Solano County
solano_ses_sf <- sesbania_3857_sf %>% 
  filter(county == 'Solano')

# subset counties based on solano
solano_sf <- ca_subset_sf %>% 
  filter(county_name == 'Solano')

ggplot() +
  geom_sf(data = solano_sf) +
  geom_sf(data = solano_ses_sf, color = 'red')
```

```{r}

tmap_mode(mode = 'view')

tm_shape(ca_subset_sf) +
  tm_fill('land_area', palette = 'BuGn') +
  tm_borders(col = 'black') +
  tm_shape(sesbania_3857_sf) +
  tm_dots(col = 'red')
```

# Part 2: Kriging

```{r}

# read in CA map and groundwater depth data
ca_count_sf <- read_sf(here("data/ca_counties"),
                       layer = 'CA_Counties_TIGER2016') %>% 
  janitor::clean_names() %>% 
  select(name)

# check projectns
st_crs(ca_count_sf)

# read in gw data
ca_dgw_sf <- read_sf(here("data", "ca_dgw",
                          "F2013_DBGS_Points_20150720_093252.gpkg")) %>% 
  janitor::clean_names()

st_crs(ca_dgw_sf)

ca_dgw_sf <- st_transform(ca_dgw_sf, st_crs(ca_count_sf))

ggplot() +
  geom_sf(data = ca_count_sf) +
  geom_sf(data = ca_dgw_sf, aes(color = dgbs))
```

```{r}

tmap_mode(mode = 'view')

tm_shape(ca_dgw_sf) + 
  tm_dots('dgbs', palette = 'Blues')

# tmaptools::palette_explorer()  ## need 'shinyjs' package
```

### Clip to San Joaquin County

```{r}

sj_county_sf <- ca_count_sf %>% 
  filter(name == 'San Joaquin')

# now keep obs for groundwater depth in SJ County

sj_depth_sf <- ca_dgw_sf %>% 
  st_intersection(sj_county_sf)

ggplot() +
  geom_sf(data = sj_county_sf) +
  geom_sf(data = sj_depth_sf, aes(color = dgbs)) +
  theme_void() +
  scale_color_viridis_c()
```

### Check for duplicate locations

```{r}

well_duplicates <- sj_depth_sf %>% 
  janitor::get_dupes(latitude, longitude)

sj_depth_sf <- sj_depth_sf %>% 
  filter(!local_well %in% well_duplicates$local_well)
```

### Create a variogram!

```{r}

sj_dgw_vgm <- gstat::variogram(dgbs ~ 1, data = sj_depth_sf)
# calculates a mean for sj county for ordinary kriging

plot(sj_dgw_vgm)
```

```{r}

# create variogram model

sj_dgw_vgm_fit <- fit.variogram(sj_dgw_vgm,
                               model = vgm(model = 'Gau',
                                           nugget = 20,
                                           sill = 3000,
                                           range = 32000))

plot(sj_dgw_vgm, sj_dgw_vgm_fit)

attr(sj_dgw_vgm_fit, 'SSErr')
```

### Spatial Kriging

```{r}

sj_grid <- st_bbox(sj_county_sf) %>% 
  st_as_stars(dx = 500, dy = 500) %>% 
  st_set_crs(st_crs(sj_county_sf)) %>% 
  st_crop(sj_county_sf)

plot(sj_grid)

# perform ordinary kriging

sj_dgw_krige <- krige(dgbs ~ 1, sj_depth_sf, sj_grid,
                      model = sj_dgw_vgm_fit)

plot(sj_dgw_krige)

krige_df <- as.data.frame(sj_dgw_krige) %>% 
  st_as_sf(coords = c('x', 'y')) %>% 
  drop_na(var1.pred)

ggplot(data = krige_df) +
  geom_sf(aes(color = var1.pred)) +
  scale_color_gradient(low = 'purple', high = 'yellow') +
  theme_void() +
  labs(color = 'Predicted\nGroundwater\nDepth (m)')
```









